---
title: "tests"
authors: "Pau Bosch & Bernat Comas"
output: html_document
---

Firstly, we load the libraries

```{r setup, include=FALSE}
library(TSP)
library(GA)
library(tidyverse) # Data manipulation
```

## Fitness function

Our fitness function will calculate the negative total distance of the path indicated by the chromosome, because we use GAs to minimize the problem.

```{r cars}
fitness_function <- function(route, dist_matrix) {
  # Add start node to the route (We are not using it)
  route <- c(route, route[1])

  total_distance <- 0
  for (i in 1:(length(route) - 1)) {
    total_distance <- total_distance + dist_matrix[route[i], route[i+1]]
  }
  return(-total_distance)  # Minimize total distance, so return negative
}
```

## Crossover functions

#### Permutation order crossover

```{r pressure, echo=FALSE}
order_crossover <- function(ga_info, parents) {

  id_parent1 <- parents[1]
  id_parent2 <- parents[2]
  
  parent1 <- ga_info@population[id_parent1, ]
  parent2 <- ga_info@population[id_parent2, ]

  n <- length(parent1)

  # Randomly select two crossover points
  crossover_points <- sort(sample(1:n, 2))
  start <- crossover_points[1]
  end <- crossover_points[2]

  # Create the child with NA values
  child1 <- rep(NA, n)
  child2 <- rep(NA, n)

  # Fill the child with the segment from the first parent
  child1[start:end] <- parent1[start:end]
  child2[1:start] <- parent1[1:start]
  child2[end:n] <- parent1[end:n] 

  # Fill the remaining positions with the genes from the second parent
  
  p2_index_ch1 <- 1
  p2_index_ch2 <- 1
  for (i in 1:n) {
    if (is.na(child1[i])) {
      while (parent2[p2_index_ch1] %in% child1) {
        p2_index_ch1 <- p2_index_ch1 + 1
      }
      child1[i] <- parent2[p2_index_ch1]
    }
    
    if (is.na(child2[i])) {
      while (parent2[p2_index_ch2] %in% child2) {
        p2_index_ch2  <- p2_index_ch2 + 1
      }
      child2[i] <- parent2[p2_index_ch2]
    }
  }

  children <- rbind(child1, child2) 
  fitness <- rep(NA, 2) # They have no fintess value sine they are different from the parents
  
  return(list(children = children, fitness = fitness))
}


```

#### Partially Matched Crossover

```{r pressure, echo=FALSE}
partially_matched_crossover <- function(ga_info, parents) {
  id_parent1 <- parents[1]
  id_parent2 <- parents[2]
  
  parent1 <- ga_info@population[id_parent1, ]
  parent2 <- ga_info@population[id_parent2, ]

  n <- length(parent1)

  # Step 1: Select two crossover points
  crossover_points <- sort(sample(1:n, 2))
  start <- crossover_points[1]
  end <- crossover_points[2]

  # Step 2: Initialize children with NA values
  child1 <- rep(NA, n)
  child2 <- rep(NA, n)

  # Step 3: Copy crossover segment from each parent
  child1[start:end] <- parent1[start:end]
  child2[start:end] <- parent2[start:end]

  # Step 4: Fill in genes from parent2 to child1 and parent1 to child2 outside the crossover segment
  for (i in 1:n) {
    # Fill for child1 from parent2
    if (is.na(child1[i])) {
      gene <- parent2[i]
      while (gene %in% child1[start:end]) {
        gene <- parent2[which(parent1 == gene)]
      }
      child1[i] <- gene
    }
    # Fill for child2 from parent1
    if (is.na(child2[i])) {
      gene <- parent1[i]
      while (gene %in% child2[start:end]) {
        gene <- parent1[which(parent2 == gene)]
      }
      child2[i] <- gene
    }
  }

  # Combine children into a matrix and initialize fitness
  children <- rbind(child1, child2) 
  fitness <- rep(NA, 2) # Fitness placeholder
  
  return(list(children = children, fitness = fitness))
}

```

#### Cycle Crossover

```{r pressure, echo=FALSE}
cycle_crossover <- function(ga_info, parents) {
  id_parent1 <- parents[1]
  id_parent2 <- parents[2]

  # Retrieve parents from the GA population
  parent1 <- ga_info@population[id_parent1, ]
  parent2 <- ga_info@population[id_parent2, ]

  n <- length(parent1)

  # Initialize children with NA values
  child1 <- rep(NA, n)
  child2 <- rep(NA, n)

  # Track cycles and fill each child based on the cycle
  cycle <- 1
  remaining_positions <- 1:n

  while (length(remaining_positions) > 0) {
    # Start at the first unassigned position
    start <- remaining_positions[1]
    pos <- start
    cycle_positions <- integer(0)

    # Follow the cycle until we return to the start position
    repeat {
      cycle_positions <- c(cycle_positions, pos)
      gene <- parent1[pos]
      pos <- which(parent2 == gene)
      if (pos == start) break
    }

    # Copy genes to children based on the cycle index
    if (cycle %% 2 == 1) {
      # Odd cycles go directly to child1 from parent1, and child2 from parent2
      child1[cycle_positions] <- parent1[cycle_positions]
      child2[cycle_positions] <- parent2[cycle_positions]
    } else {
      # Even cycles go to child1 from parent2, and child2 from parent1
      child1[cycle_positions] <- parent2[cycle_positions]
      child2[cycle_positions] <- parent1[cycle_positions]
    }

    # Update remaining positions and cycle count
    remaining_positions <- setdiff(remaining_positions, cycle_positions)
    cycle <- cycle + 1
  }

  # Combine children into a matrix and initialize fitness
  children <- rbind(child1, child2)
  fitness <- rep(NA, 2) # Placeholder for fitness values
  
  return(list(children = children, fitness = fitness))
}
```

#### Maximal Preservative Crossover

```{r}
maximal_preservative_crossover <- function(ga_info, parents) {

  # Identify parents by their indices in the population
  id_parent1 <- parents[1]
  id_parent2 <- parents[2]
  
  # Retrieve the parent chromosomes from the population
  parent1 <- ga_info@population[id_parent1, ]
  parent2 <- ga_info@population[id_parent2, ]
  n <- length(parent1)
  
  # Initialize children as NA to be filled later
  child1 <- rep(NA, n)
  child2 <- rep(NA, n)
  
  # Step 1: Select a random contiguous segment in Parent 1
  segment_start <- sample(1:(n-1), 1)
  segment_end <- sample((segment_start + 1):n, 1)
  
  # Copy the selected segment from Parent 1 to Child 1
  child1[segment_start:segment_end] <- parent1[segment_start:segment_end]
  child2[segment_start:segment_end] <- parent2[segment_start:segment_end]
  
  # Function to fill remaining positions from other parent
  fill_remaining_positions <- function(child, parent, segment) {
    current_pos <- 1
    for (i in 1:n) {
      if (!(parent[i] %in% segment)) {
        while (!is.na(child[current_pos])) {
          current_pos <- current_pos + 1
        }
        child[current_pos] <- parent[i]
      }
    }
    return(child)
  }
  
  # Step 2: Fill remaining cities from Parent 2 for Child 1 and from Parent 1 for Child 2
  child1 <- fill_remaining_positions(child1, parent2, parent1[segment_start:segment_end])
  child2 <- fill_remaining_positions(child2, parent1, parent2[segment_start:segment_end])

  # Final validation to ensure no duplicates and correct length
  if (length(unique(child1)) != n || any(is.na(child1))) {
    print("Invalid child1 detected. Returning parents.")
    return(list(children = rbind(parent1, parent2), fitness = rep(NA, 2)))
  }
  
  if (length(unique(child2)) != n || any(is.na(child2))) {
    print("Invalid child2 detected. Returning parents.")
    return(list(children = rbind(parent1, parent2), fitness = rep(NA, 2)))
  }
  
  return(list(children = rbind(child1, child2), fitness = rep(NA, 2)))
}

```

## Mutation functions

#### Swap mutation permutation

```{r}
swap_mutation <- function(ga_info, chromosome_id) {

  chromosome <- ga_info@population[chromosome_id, ]
  
  n <- length(chromosome)
  
  # Randomly select two positions to swap
  positions <- sample(1:n, 2)
  
  # Swap the genes at the selected positions
  temp <- chromosome[positions[1]]
  chromosome[positions[1]] <- chromosome[positions[2]]
  chromosome[positions[2]] <- temp

  if (length(chromosome) != 52){
    print(chromosome)
  }

return(chromosome)
}
```

#### Inverse mutation

```{r}
inverse_mutation <- function(ga_info, chromosome_id) {
  
  # Get the chromosome to mutate
  chromosome <- ga_info@population[chromosome_id, ]
  
  # Get the number of genes in the chromosome
  n <- length(chromosome)
  
  # Randomly select two positions for the inversion
  positions <- sort(sample(1:n, 2))
  
  # Reverse the sequence between the selected positions
  chromosome[positions[1]:positions[2]] <- rev(chromosome[positions[1]:positions[2]])
  
  return(chromosome)
}

```

#### Insert mutation:

```{r}
insert_mutation <- function(ga_info, chromosome_id) {
  
  # Get the chromosome to mutate
  chromosome <- ga_info@population[chromosome_id, ]
  
  # Get the number of genes in the chromosome
  n <- length(chromosome)
  
  # Randomly select two positions: one to remove from and one to insert into
  positions <- sample(1:n, 2)
  pos_from <- positions[1]
  pos_to <- positions[2]
  
  # Extract the gene at pos_from
  gene <- chromosome[pos_from]
  
  # Remove the gene at pos_from
  chromosome <- chromosome[-pos_from]
  
  # Insert the gene at pos_to
  chromosome <- append(chromosome, gene, after = pos_to - 1)
  
  return(chromosome)
}

```

#### Scramble mutation

```{r}
scramble_mutation <- function(ga_info, chromosome_id) {
  
  # Get the chromosome to mutate
  chromosome <- ga_info@population[chromosome_id, ]
  
  # Get the number of genes in the chromosome
  n <- length(chromosome)
  
  # Randomly select two positions for the inversion
  positions <- sort(sample(1:n, 2))
  
  # Reverse the sequence between the selected positions
  chromosome[positions[1]:positions[2]] <- sample(chromosome[positions[1]:positions[2]])
  
  return(chromosome)
}
```

## GA algorithm

Run the functions:

```{r}
calculate_distance_matrix <- function(coords) {
  num_cities <- nrow(coords)
  
  # Matrix to store the distances
  dist_matrix <- matrix(0, nrow = num_cities, ncol = num_cities)
  
  # Calculate the distances between every pair of cities
  for (i in 1:num_cities) {
    for (j in 1:num_cities) {
      # Calculate Euclidean distance between city i and city j
      dist_matrix[i, j] <- sqrt((coords[i, 1] - coords[j, 1])^2 + (coords[i, 2] - coords[j, 2])^2)
    }
  }
  
  return(dist_matrix)
}

plot_tsp_route <- function(coords, route) {
  plot(coords, type = "n", main = "Optimized TSP Route", xlab = "X Coordinate", ylab = "Y Coordinate")
  points(coords, pch = 16, col = "blue")  # Plot cities
  text(coords, labels = 1:nrow(coords), pos = 4, cex = 0.8, col = "blue")  # Label cities

  for (i in 1:(length(route) - 1)) {
    city_from <- route[i]
    city_to <- route[i + 1]
    lines(coords[c(city_from, city_to), ], col = "red", lwd = 2)
  }
}
```

```{r}
ins <- read_TSPLIB("instances/berlin52.tsp")

print(ins)
```

```{r}
dist_matrix <- calculate_distance_matrix(ins)

print(dim(dist_matrix))
```

```{r}
population_size <- 500
max_generations <- 50000
crossover_prob <- 0.6
mutation_prob <- 0.2
elitism <- base::max(1, round(population_size*0.2))
convergence_criteria <- 100  # Number of generations with no improvement on the best fitness.


ga_result <- ga(type = "permutation",
    fitness = function(route) fitness_function(route, dist_matrix),
    lower = 1, upper = nrow(dist_matrix),
    maxiter = max_generations,
    popSize = population_size,  # Population size
    pmutation = mutation_prob,  # Mutation probability
    pcrossover = crossover_prob,  # Crossover probability
    crossover = order_crossover,
    mutation = inverse_mutation,
    keepBest = TRUE,
    elitism = elitism,
    run = convergence_criteria
  )

summary(ga_result)
plot(ga_result)

```

```{r}
# Plot
best_route <- c(ga_result@solution, ga_result@solution[1])  # Loop back to the start city
plot_tsp_route(ins, best_route)
```

## Perform grid search

```{r}
# Define the parameter grid
crossover_operators <- list(
  "order" = order_crossover,
  "partially_matched" = partially_matched_crossover,
  "maximal_preservative" = maximal_preservative_crossover,
  "cycle" = cycle_crossover
  )

mutation_operators <- list(
  "swap_mutation" = swap_mutation,
  "inverse_mutation" = inverse_mutation,
  "insert_mutation" = insert_mutation,
  "scramble_mutation" = scramble_mutation)

elitism <- c(0.1, 0.2, 0.3)
mutation_probs <- c(0.1, 0.2, 0.3)
crossover_probs <- c(0.6, 0.8, 0.9) 
population_size <- 500

# Prepare to store results
results <- list()
output_dir <- "./GA_results"
dir.create(output_dir, showWarnings = FALSE)

# Define a function to run the GA with specific parameters
run_ga_with_params <- function(crossover_op, mutation_op, elitism_val, mut_prob, cross_prob, dist_matrix, population_size) {
  crossover_wrapper <- function(ga_info, parents) {
    return(crossover_op(ga_info, parents))
  }
  
  ga <- ga(
    type = "permutation",
    fitness = function(route) fitness_function(route, dist_matrix),
    lower = 1, upper = nrow(dist_matrix),
    popSize = population_size,
    maxiter = 500,
    elitism = base::max(1, round(population_size * elitism_val)),
    pmutation = mut_prob,
    pcrossover = cross_prob,
    crossover = crossover_wrapper,
    mutation = mutation_op,
    keepBest = TRUE,
    monitor = NULL
  )
  return(ga)
}

generate_comb_name <- function(crossover_name, mutation_name, elitism_val, mut_prob, cross_prob) {
  paste0(
    crossover_name, 
    "_", mutation_name, 
    "_elitism_", elitism_val, 
    "_mutprob_", mut_prob, 
    "_crossprob_", cross_prob
  )
}


output_dir <- "./GA_results"
dir.create(output_dir, showWarnings = FALSE)
csv_file <- paste0(output_dir, "/all_results.csv")

write.csv(data.frame(crossover = character(), mutation = character(), elitism = numeric(),
                     mutation_prob = numeric(), crossover_prob = numeric(), best_fitness = numeric(), best_solution = character()),
          file = csv_file, row.names = FALSE, append = FALSE)

run_and_save_ga <- function(crossover_name, mutation_name, elitism_val, mut_prob, cross_prob, dist_matrix, population_size) {
  crossover_op <- crossover_operators[[crossover_name]]
  mutation_op <- mutation_operators[[mutation_name]]
  
  comb_name <- generate_comb_name(crossover_name, mutation_name, elitism_val, mut_prob, cross_prob)
  
  ga_result <- run_ga_with_params(crossover_op, mutation_op, elitism_val, mut_prob, cross_prob, dist_matrix, population_size)
  
  fitness_history <- ga_result@fitness
  
  result_data <- data.frame(
    crossover = crossover_name,
    mutation = mutation_name,
    elitism = elitism_val,
    mutation_prob = mut_prob,
    crossover_prob = cross_prob,
    best_fitness = max(ga_result@fitness),
    best_solution = paste(ga_result@solution, collapse = "|")
  )
  
  print(paste("Saving result for combination:", comb_name))
  
  write.table(result_data, file = csv_file, row.names = FALSE, col.names = FALSE, sep = ",", append = TRUE)

  png(file = paste0(output_dir, "/learning_curve_", comb_name, ".png"))
  plot(ga_result)
  dev.off()

  best_route <- c(ga_result@solution, ga_result@solution[1])  # Loop back to the start
  png(file = paste0(output_dir, "/solution_curve_", comb_name, ".png"))
  plot_tsp_route(ins, best_route)
  dev.off()
}

for (crossover_name in names(crossover_operators)) {
  for (mutation_name in names(mutation_operators)) {
    for (elitism_val in elitism) {
      for (mut_prob in mutation_probs) {
        for (cross_prob in crossover_probs) {
          print(paste("Running GA with crossover:", crossover_name, 
                      "mutation:", mutation_name, 
                      "elitism:", elitism_val, 
                      "mutation_prob:", mut_prob, 
                      "crossover_prob:", cross_prob))
          run_and_save_ga(crossover_name, mutation_name, elitism_val, mut_prob, cross_prob, dist_matrix, population_size)
        }
      }
    }
  }
}

```

## Islands Genetic Algorithm

```{r}
library(parallel)
library(doParallel)
```

```{r}
population_size <- 500
max_generations <- 800
crossover_prob <- 0.8
mutation_prob <- 0.07
elitism <- 100

ga_result <- gaisl(type = "permutation",
    fitness = function(route) fitness_function(route, dist_matrix),
    lower = 1, upper = nrow(dist_matrix),
    popSize = population_size,  # Population size
    maxiter = max_generations,  # Number of generations
    pmutation = mutation_prob,  # Mutation probability
    pcrossover = crossover_prob,  # Crossover probability
    crossover = order_crossover,
    mutation = inverse_mutation,
    elitism = elitism
  )

summary(ga_result)
```

# Optimal Result Extraction

```{r}
# Check what is the optimal fitness
opt_bayg29 <- c(1, 28, 6, 12, 9, 26, 3, 29, 5, 21, 2, 20, 10, 4, 15, 18, 14, 17, 22, 11, 19, 25, 7, 23, 8, 27, 16, 13, 24)

opt_berlin52 <- c(1, 49, 32, 45, 19, 41, 8, 9, 10, 43, 33, 51, 11, 52, 14, 13, 47, 26, 27, 28, 12, 25, 4, 6, 15, 5, 24, 48, 38, 37, 40, 39, 36, 35, 34, 44, 46, 16, 29, 50, 20, 23, 30, 2, 7, 42, 21, 17, 3, 18, 31, 22)

opt_tsp225 <- c(1, 200, 198, 197, 195, 194, 218, 193, 196, 192, 191, 205, 189, 27, 188, 187, 186, 185, 184, 182, 181, 180, 179, 178, 177, 176, 174, 173, 172, 171, 170, 169, 168, 212, 214, 167, 166, 165, 164, 213, 158, 163, 162, 161, 160, 159, 157, 156, 155, 154, 153, 152, 151, 150, 149, 148, 147, 146, 145, 144, 143, 201, 142, 141, 140, 139, 138, 137, 136, 183, 135, 134, 215, 132, 131, 211, 130, 222, 129, 128, 127, 126, 125, 124, 123, 122, 121, 175, 120, 119, 118, 117, 116, 223, 115, 114, 113, 112, 111, 110, 109, 108, 107, 106, 105, 220, 104, 103, 102, 101, 100, 99, 98, 97, 96, 95, 209, 94, 93, 92, 91, 90, 89, 88, 87, 210, 86, 85, 84, 83, 82, 221, 81, 80, 79, 78, 77, 217, 219, 216, 76, 75, 74, 73, 72, 71, 70, 69, 68, 67, 66, 65, 64, 63, 62, 61, 60, 59, 58, 57, 56, 55, 54, 53, 52, 50, 51, 49, 207, 2, 47, 225, 190, 133, 199, 224, 48, 45, 46, 44, 43, 42, 41, 40, 39, 38, 37, 36, 34, 33, 35, 32, 31, 206, 202, 30, 29, 28, 204, 26, 25, 208, 24, 23, 22, 21, 20, 203, 19, 18, 17, 16, 15, 14, 13, 12, 11, 10, 9, 8, 7, 6, 5, 4, 3)

print("BAYG29")
print(length(opt_bayg29))
ins <- read_TSPLIB("instances/bayg29.tsp")
dist_matrix <- calculate_distance_matrix(ins)
print(fitness_function(opt_bayg29,dist_matrix))
plot_tsp_route(ins, opt_bayg29)

print("BERLIN52")
print(length(opt_berlin52))
ins <- read_TSPLIB("instances/berlin52.tsp")
dist_matrix <- calculate_distance_matrix(ins)
print(fitness_function(opt_berlin52,dist_matrix))
plot_tsp_route(ins, opt_berlin52)

print("TSP225")
print(length(opt_tsp225))
ins <- read_TSPLIB("instances/tsp225.tsp")
dist_matrix <- calculate_distance_matrix(ins)
print(fitness_function(opt_tsp225,dist_matrix))
plot_tsp_route(ins, opt_tsp225)
```
