---
title: "tests"
output: html_document
---

Firstly, we load the libraries

```{r setup, include=FALSE}
library(TSP)
library(GA)
library(tidyverse) # Data manipulation
```

## Fitness function

Our fitness function will calculate the negative total distance of the path indicated by the chromosome, because we use GAs to minimize the problem.

```{r cars}
fitness_function <- function(route, dist_matrix) {
  # Add start node to the route (We are not using it)
  route <- c(route, route[1])

  total_distance <- 0
  for (i in 1:(length(route) - 1)) {
    total_distance <- total_distance + dist_matrix[route[i], route[i+1]]
  }
  return(-total_distance)  # Minimize total distance, so return negative
}
```

## Crossover functions

Permutation order crossover

```{r pressure, echo=FALSE}
order_crossover <- function(ga_info, parents) {

  id_parent1 <- parents[1]
  id_parent2 <- parents[2]
  
  parent1 <- ga_info@population[id_parent1, ]
  parent2 <- ga_info@population[id_parent2, ]

  n <- length(parent1)

  # Randomly select two crossover points
  crossover_points <- sort(sample(1:n, 2))
  start <- crossover_points[1]
  end <- crossover_points[2]

  # Create the child with NA values
  child1 <- rep(NA, n)
  child2 <- rep(NA, n)

  # Fill the child with the segment from the first parent
  child1[start:end] <- parent1[start:end]
  child2[1:start] <- parent1[1:start]
  child2[end:n] <- parent1[end:n] 

  # Fill the remaining positions with the genes from the second parent
  
  p2_index_ch1 <- 1
  p2_index_ch2 <- 1
  for (i in 1:n) {
    if (is.na(child1[i])) {
      while (parent2[p2_index_ch1] %in% child1) {
        p2_index_ch1 <- p2_index_ch1 + 1
      }
      child1[i] <- parent2[p2_index_ch1]
    }
    
    if (is.na(child2[i])) {
      while (parent2[p2_index_ch2] %in% child2) {
        p2_index_ch2  <- p2_index_ch2 + 1
      }
      child2[i] <- parent2[p2_index_ch2]
    }
  }

  children <- rbind(child1, child2) 
  fitness <- rep(NA, 2) # They have no fintess value sine they are different from the parents
  
  return(list(children = children, fitness = fitness))
}


```

# Partially Matched Crossover

```{r pressure, echo=FALSE}
partially_matched_crossover <- function(ga_info, parents) {
  id_parent1 <- parents[1]
  id_parent2 <- parents[2]
  
  parent1 <- ga_info@population[id_parent1, ]
  parent2 <- ga_info@population[id_parent2, ]

  n <- length(parent1)

  # Step 1: Select two crossover points
  crossover_points <- sort(sample(1:n, 2))
  start <- crossover_points[1]
  end <- crossover_points[2]

  # Step 2: Initialize children with NA values
  child1 <- rep(NA, n)
  child2 <- rep(NA, n)

  # Step 3: Copy crossover segment from each parent
  child1[start:end] <- parent1[start:end]
  child2[start:end] <- parent2[start:end]

  # Step 4: Fill in genes from parent2 to child1 and parent1 to child2 outside the crossover segment
  for (i in 1:n) {
    # Fill for child1 from parent2
    if (is.na(child1[i])) {
      gene <- parent2[i]
      while (gene %in% child1[start:end]) {
        gene <- parent2[which(parent1 == gene)]
      }
      child1[i] <- gene
    }
    # Fill for child2 from parent1
    if (is.na(child2[i])) {
      gene <- parent1[i]
      while (gene %in% child2[start:end]) {
        gene <- parent1[which(parent2 == gene)]
      }
      child2[i] <- gene
    }
  }

  # Combine children into a matrix and initialize fitness
  children <- rbind(child1, child2) 
  fitness <- rep(NA, 2) # Fitness placeholder
  
  return(list(children = children, fitness = fitness))
}

```


# Cycle Crossover

```{r pressure, echo=FALSE}
cycle_crossover <- function(ga_info, parents) {
  id_parent1 <- parents[1]
  id_parent2 <- parents[2]

  # Retrieve parents from the GA population
  parent1 <- ga_info@population[id_parent1, ]
  parent2 <- ga_info@population[id_parent2, ]

  n <- length(parent1)

  # Initialize children with NA values
  child1 <- rep(NA, n)
  child2 <- rep(NA, n)

  # Track cycles and fill each child based on the cycle
  cycle <- 1
  remaining_positions <- 1:n

  while (length(remaining_positions) > 0) {
    # Start at the first unassigned position
    start <- remaining_positions[1]
    pos <- start
    cycle_positions <- integer(0)

    # Follow the cycle until we return to the start position
    repeat {
      cycle_positions <- c(cycle_positions, pos)
      gene <- parent1[pos]
      pos <- which(parent2 == gene)
      if (pos == start) break
    }

    # Copy genes to children based on the cycle index
    if (cycle %% 2 == 1) {
      # Odd cycles go directly to child1 from parent1, and child2 from parent2
      child1[cycle_positions] <- parent1[cycle_positions]
      child2[cycle_positions] <- parent2[cycle_positions]
    } else {
      # Even cycles go to child1 from parent2, and child2 from parent1
      child1[cycle_positions] <- parent2[cycle_positions]
      child2[cycle_positions] <- parent1[cycle_positions]
    }

    # Update remaining positions and cycle count
    remaining_positions <- setdiff(remaining_positions, cycle_positions)
    cycle <- cycle + 1
  }

  # Combine children into a matrix and initialize fitness
  children <- rbind(child1, child2)
  fitness <- rep(NA, 2) # Placeholder for fitness values
  
  return(list(children = children, fitness = fitness))
}
```

Edge Recombination Crossover (ERX) NO S UTILITZA

```{r}
edge_recombination_crossover <- function(ga_info, parents) {
  id_parent1 <- parents[1]
  id_parent2 <- parents[2]
  
  parent1 <- ga_info@population[id_parent1, ]
  parent2 <- ga_info@population[id_parent2, ]
  
  n <- length(parent1)
  
  # Step 1: Create the edge map
  create_edge_map <- function(p1, p2) {
    edge_map <- vector("list", n)
    
    for (i in 1:n) {
      neighbors <- c()
      # Add neighbors from parent1
      idx <- which(p1 == i)
      neighbors <- c(neighbors, p1[(idx - 1 - 1) %% n + 1], p1[(idx + 1 - 1) %% n + 1])
      
      # Add neighbors from parent2
      idx <- which(p2 == i)
      neighbors <- c(neighbors, p2[(idx - 1 - 1) %% n + 1], p2[(idx + 1 - 1) %% n + 1])
      
      # Remove duplicates
      edge_map[[i]] <- unique(neighbors)
    }
    
    return(edge_map)
  }
  
  edge_map <- create_edge_map(parent1, parent2)
  
  # Step 2: Create child by starting from a random city
  generate_child <- function(edge_map) {
    child <- numeric(n)
    visited <- rep(FALSE, n)
    
    # Start from a random city
    current <- sample(1:n, 1)
    child[1] <- current
    visited[current] <- TRUE
    
    for (i in 2:n) {
      # Remove the current city from its neighbors' lists
      for (j in 1:n) {
        edge_map[[j]] <- setdiff(edge_map[[j]], current)
      }
      
      # Get the neighbors of the current city
      neighbors <- edge_map[[current]]
      
      # Choose the next city
      if (length(neighbors) > 0) {
        # Select the neighbor with the fewest edges in its list
        next_city <- neighbors[which.min(sapply(neighbors, function(x) length(edge_map[[x]])))]
      } else {
        # If no neighbors left, pick an unvisited city at random
        next_city <- sample(which(!visited), 1)
      }
      
      # Add the selected city to the child and mark it as visited
      child[i] <- next_city
      visited[next_city] <- TRUE
      current <- next_city
    }
    
    return(child)
  }
  
  # Generate two children
  child1 <- generate_child(edge_map)
  child2 <- generate_child(edge_map)
  
  # Combine children into a matrix and initialize fitness
  children <- rbind(child1, child2)
  fitness <- rep(NA, 2) # Fitness placeholder
  
  return(list(children = children, fitness = fitness))
}

```


### Greedy crossover NO FUNCIONA
```{r}

get_greedy_child <- function(parent1, parent2, distance_matrix){
  n <- length(parent1)
  child <- rep(NA, n)
  child[1] <- parent1[1]

  for (i in 1:(n - 1)) {
    current_city <- child[i]
    
    pos_in_parent1 <- match(current_city, parent1)
    pos_in_parent2 <- match(current_city, parent2)
    
    next_in_parent1 <- parent1[pos_in_parent1 + 1]
    next_in_parent2 <- parent2[pos_in_parent2 + 1]
    
    dist1 <- if (!is.na(next_in_parent1) && !(next_in_parent1 %in% child)) 
               distance_matrix[current_city, next_in_parent1] 
             else Inf
    dist2 <- if (!is.na(next_in_parent2) && !(next_in_parent2 %in% child)) 
               distance_matrix[current_city, next_in_parent2] 
             else Inf
    
    if (!is.infinite(dist1) && !is.infinite(dist2)) {
      next_city <- if (dist1 < dist2) next_in_parent1 else next_in_parent2
    } else if (!is.infinite(dist1)) {
      next_city <- next_in_parent1
    } else if (!is.infinite(dist2)) {
      next_city <- next_in_parent2
    } else {
      unvisited_cities <- setdiff(parent1, child)
      
      if (length(unvisited_cities) > 0) {
        next_city <- unvisited_cities[1]
      } else {
        next_city <- NA
        print("No available city")
      }
    }
    
    child[i + 1] <- next_city
  }
  
  # Final validation to ensure no duplicates and correct length
  if (length(child) != 52 || any(is.na(child)) || length(child) != n) {
    print("Invalid child1 detected:")
    print(child)
    print("Returning parents due to invalid child1.")
    return(parent1)
  }
  
  return(child)

}

adaptive_greedy_subpath_crossover <- function(ga_info, parents, distance_matrix) {

  id_parent1 <- parents[1]
  id_parent2 <- parents[2]
  
  parent1 <- ga_info@population[id_parent1, ]
  parent2 <- ga_info@population[id_parent2, ]
  n <- length(parent1)
  
  child1 <- get_greedy_child(parent1, parent2, distance_matrix)
  child2 <- get_greedy_child(parent2, parent1, distance_matrix)
  
  children <- rbind(child1, child2) 
  fitness <- rep(NA, 2)
  
  return(list(children = children, fitness = fitness))
}


```

### Maximal Preservative Crossover

```{r}
maximal_preservative_crossover <- function(ga_info, parents) {

  # Identify parents by their indices in the population
  id_parent1 <- parents[1]
  id_parent2 <- parents[2]
  
  # Retrieve the parent chromosomes from the population
  parent1 <- ga_info@population[id_parent1, ]
  parent2 <- ga_info@population[id_parent2, ]
  n <- length(parent1)
  
  # Initialize children as NA to be filled later
  child1 <- rep(NA, n)
  child2 <- rep(NA, n)
  
  # Step 1: Select a random contiguous segment in Parent 1
  segment_start <- sample(1:(n-1), 1)
  segment_end <- sample((segment_start + 1):n, 1)
  
  # Copy the selected segment from Parent 1 to Child 1
  child1[segment_start:segment_end] <- parent1[segment_start:segment_end]
  child2[segment_start:segment_end] <- parent2[segment_start:segment_end]
  
  # Function to fill remaining positions from other parent
  fill_remaining_positions <- function(child, parent, segment) {
    current_pos <- 1
    for (i in 1:n) {
      if (!(parent[i] %in% segment)) {
        while (!is.na(child[current_pos])) {
          current_pos <- current_pos + 1
        }
        child[current_pos] <- parent[i]
      }
    }
    return(child)
  }
  
  # Step 2: Fill remaining cities from Parent 2 for Child 1 and from Parent 1 for Child 2
  child1 <- fill_remaining_positions(child1, parent2, parent1[segment_start:segment_end])
  child2 <- fill_remaining_positions(child2, parent1, parent2[segment_start:segment_end])

  # Final validation to ensure no duplicates and correct length
  if (length(unique(child1)) != n || any(is.na(child1))) {
    print("Invalid child1 detected. Returning parents.")
    return(list(children = rbind(parent1, parent2), fitness = rep(NA, 2)))
  }
  
  if (length(unique(child2)) != n || any(is.na(child2))) {
    print("Invalid child2 detected. Returning parents.")
    return(list(children = rbind(parent1, parent2), fitness = rep(NA, 2)))
  }
  
  return(list(children = rbind(child1, child2), fitness = rep(NA, 2)))
}

```


## Mutation functions

Swap mutation permutation

```{r}
swap_mutation <- function(ga_info, chromosome_id) {

  chromosome <- ga_info@population[chromosome_id, ]
  
  n <- length(chromosome)
  
  # Randomly select two positions to swap
  positions <- sample(1:n, 2)
  
  # Swap the genes at the selected positions
  temp <- chromosome[positions[1]]
  chromosome[positions[1]] <- chromosome[positions[2]]
  chromosome[positions[2]] <- temp

  if (length(chromosome) != 52){
    print(chromosome)
  }

return(chromosome)
}
```

Inverse mutation

```{r}
inverse_mutation <- function(ga_info, chromosome_id) {
  
  # Get the chromosome to mutate
  chromosome <- ga_info@population[chromosome_id, ]
  
  # Get the number of genes in the chromosome
  n <- length(chromosome)
  
  # Randomly select two positions for the inversion
  positions <- sort(sample(1:n, 2))
  
  # Reverse the sequence between the selected positions
  chromosome[positions[1]:positions[2]] <- rev(chromosome[positions[1]:positions[2]])
  
  return(chromosome)
}

```

Insert mutation:

```{r}
insert_mutation <- function(ga_info, chromosome_id) {
  
  # Get the chromosome to mutate
  chromosome <- ga_info@population[chromosome_id, ]
  
  # Get the number of genes in the chromosome
  n <- length(chromosome)
  
  # Randomly select two positions: one to remove from and one to insert into
  positions <- sample(1:n, 2)
  pos_from <- positions[1]
  pos_to <- positions[2]
  
  # Extract the gene at pos_from
  gene <- chromosome[pos_from]
  
  # Remove the gene at pos_from
  chromosome <- chromosome[-pos_from]
  
  # Insert the gene at pos_to
  chromosome <- append(chromosome, gene, after = pos_to - 1)
  
  return(chromosome)
}

```

Scramble mutation 

```{r}
scramble_mutation <- function(ga_info, chromosome_id) {
  
  # Get the chromosome to mutate
  chromosome <- ga_info@population[chromosome_id, ]
  
  # Get the number of genes in the chromosome
  n <- length(chromosome)
  
  # Randomly select two positions for the inversion
  positions <- sort(sample(1:n, 2))
  
  # Reverse the sequence between the selected positions
  chromosome[positions[1]:positions[2]] <- sample(chromosome[positions[1]:positions[2]])
  
  return(chromosome)
}
```

## GA algorithm

Run the functions:

```{r}
calculate_distance_matrix <- function(coords) {
  num_cities <- nrow(coords)
  
  # Matrix to store the distances
  dist_matrix <- matrix(0, nrow = num_cities, ncol = num_cities)
  
  # Calculate the distances between every pair of cities
  for (i in 1:num_cities) {
    for (j in 1:num_cities) {
      # Calculate Euclidean distance between city i and city j
      dist_matrix[i, j] <- sqrt((coords[i, 1] - coords[j, 1])^2 + (coords[i, 2] - coords[j, 2])^2)
    }
  }
  
  return(dist_matrix)
}

plot_tsp_route <- function(coords, route) {
  plot(coords, type = "n", main = "Optimized TSP Route", xlab = "X Coordinate", ylab = "Y Coordinate")
  points(coords, pch = 16, col = "blue")  # Plot cities
  text(coords, labels = 1:nrow(coords), pos = 4, cex = 0.8, col = "blue")  # Label cities

# Draw the path
  for (i in 1:(length(route) - 1)) {
    city_from <- route[i]
    city_to <- route[i + 1]
    lines(coords[c(city_from, city_to), ], col = "red", lwd = 2)
  }
}
```

```{r}
ins <- read_TSPLIB("instances/berlin52.tsp")

print(ins)
```

```{r}
dist_matrix <- calculate_distance_matrix(ins)

print(dim(dist_matrix))
```

```{r}
population_size <- 500
max_generations <- 500
crossover_prob <- 0.8
mutation_prob <- 0.1
elitism <- base::max(1, round(population_size*0.2))

ga_result <- ga(type = "permutation",  # Chromosomes are permutations (tours)
    fitness = function(route) fitness_function(route, dist_matrix),
    lower = 1, upper = nrow(dist_matrix),  # Cities are indexed from 1 to num_cities
    popSize = population_size,  # Population size
    maxiter = max_generations,  # Number of generations
    pmutation = mutation_prob,  # Mutation probability
    pcrossover = crossover_prob,  # Crossover probability
    crossover = function(ga_info, parents) random_crossover_selection(ga_info, parents, dist_matrix),  # Pass dist_matrix to crossover
    mutation = inverse_mutation, #gaperm_swMutation  # Mutation function swap_mutation inverse_mutation scramble_mutation insert_mutation
    keepBest = TRUE,
    elitism = elitism
  )

summary(ga_result)
plot(ga_result)

```

```{r}
# Plot
best_route <- c(ga_result@solution, ga_result@solution[1])  # Loop back to the start city
plot_tsp_route(ins, best_route)
```

## Perform grid search

```{r}
# Define the parameter grid
crossover_operators <- list(
  "order" = order_crossover,
  "partially_matched" = partially_matched_crossover,
  "maximal_preservative" = maximal_preservative_crossover,
  "cycle" = cycle_crossover
  )

mutation_operators <- list(
  "swap_mutation" = swap_mutation,
  "inverse_mutation" = inverse_mutation,
  "insert_mutation" = insert_mutation,
  "scramble_mutation" = scramble_mutation)

elitism <- c(0.1, 0.2, 0.3)
mutation_probs <- c(0.1, 0.3, 0.2)
crossover_probs <- c(0.6, 0.9, 0.8) 
population_size <- 500

# Prepare to store results
results <- list()
output_dir <- "./GA_results"
dir.create(output_dir, showWarnings = FALSE)

# Define a function to run the GA with specific parameters
run_ga_with_params <- function(crossover_op, mutation_op, elitism_val, mut_prob, cross_prob, dist_matrix, population_size) {
  # Define a wrapper for the crossover operation to handle distance_matrix when needed
  crossover_wrapper <- function(ga_info, parents) {
    if (identical(crossover_op, adaptive_greedy_subpath_crossover)) {
      return(crossover_op(ga_info, parents, dist_matrix))
    } else {
      return(crossover_op(ga_info, parents))
    }
  }
  
  # Run the genetic algorithm with the crossover wrapper
  ga <- ga(
    type = "permutation",
    fitness = function(route) fitness_function(route, dist_matrix),
    lower = 1, upper = nrow(dist_matrix),
    popSize = population_size,
    maxiter = 500,
    elitism = base::max(1, round(population_size * elitism_val)),
    pmutation = mut_prob,
    pcrossover = cross_prob,
    crossover = crossover_wrapper,
    mutation = mutation_op,
    monitor = NULL
  )
  return(ga)
}

generate_comb_name <- function(crossover_name, mutation_name, elitism_val, mut_prob, cross_prob) {
  paste0(
    crossover_name, 
    "_", mutation_name, 
    "_elitism_", elitism_val, 
    "_mutprob_", mut_prob, 
    "_crossprob_", cross_prob
  )
}


output_dir <- "./GA_results"
dir.create(output_dir, showWarnings = FALSE)
csv_file <- paste0(output_dir, "/all_results.csv")

# Write the CSV header only once
write.csv(data.frame(crossover = character(), mutation = character(), elitism = numeric(),
                     mutation_prob = numeric(), crossover_prob = numeric(), best_fitness = numeric(), best_solution = character()),
          file = csv_file, row.names = FALSE, append = FALSE)

# Modify function to append results to the CSV
run_and_save_ga <- function(crossover_name, mutation_name, elitism_val, mut_prob, cross_prob, dist_matrix, population_size) {
  crossover_op <- crossover_operators[[crossover_name]]
  mutation_op <- mutation_operators[[mutation_name]]
  
  comb_name <- generate_comb_name(crossover_name, mutation_name, elitism_val, mut_prob, cross_prob)
  
  ga_result <- run_ga_with_params(crossover_op, mutation_op, elitism_val, mut_prob, cross_prob, dist_matrix, population_size)
  
  fitness_history <- tryCatch({
    ga_result@fitness
  }, error = function(e) {
    message("Fitness history not directly available; using summary.")
    summary(ga_result)$fitness
  })
  
  result_data <- data.frame(
    crossover = crossover_name,
    mutation = mutation_name,
    elitism = elitism_val,
    mutation_prob = mut_prob,
    crossover_prob = cross_prob,
    best_fitness = max(ga_result@fitness),
    best_solution = paste(ga_result@solution, collapse = "|")
  )
  
  print(paste("Saving result for combination:", comb_name))
  
  write.table(result_data, file = csv_file, row.names = FALSE, col.names = FALSE, sep = ",", append = TRUE)

  png(file = paste0(output_dir, "/learning_curve_", comb_name, ".png"))
  plot(ga_result)
  dev.off()

  best_route <- c(ga_result@solution, ga_result@solution[1])  # Loop back to the start
  png(file = paste0(output_dir, "/solution_curve_", comb_name, ".png"))
  plot_tsp_route(ins, best_route)
  dev.off()
}

for (crossover_name in names(crossover_operators)) {
  for (mutation_name in names(mutation_operators)) {
    for (elitism_val in elitism) {
      for (mut_prob in mutation_probs) {
        for (cross_prob in crossover_probs) {
          print(paste("Running GA with crossover:", crossover_name, 
                      "mutation:", mutation_name, 
                      "elitism:", elitism_val, 
                      "mutation_prob:", mut_prob, 
                      "crossover_prob:", cross_prob))
          run_and_save_ga(crossover_name, mutation_name, elitism_val, mut_prob, cross_prob, dist_matrix, population_size)
        }
      }
    }
  }
}

```

```{r}
# Convert results to data frame and save as CSV
results_df <- do.call(rbind, lapply(results, as.data.frame))
write.csv(results_df, file = paste0(output_dir, "/grid_search_results.csv"), row.names = FALSE)
```


```{r}
library(parallel)
library(doParallel)
```


ISLANDS GENETIC ALGORITHM

```{r}
population_size <- 500
max_generations <- 800
crossover_prob <- 0.8
mutation_prob <- 0.07
elitism <- 100

ga_result <- gaisl(type = "permutation",  # Chromosomes are permutations (tours)
    fitness = function(route) fitness_function(route, dist_matrix),
    lower = 1, upper = nrow(dist_matrix),  # Cities are indexed from 1 to num_cities
    popSize = population_size,  # Population size
    maxiter = max_generations,  # Number of generations
    pmutation = mutation_prob,  # Mutation probability
    pcrossover = crossover_prob,  # Crossover probability
    crossover = order_crossover, #gaperm_oxCrossover,  # Crossover function: cycle_crossover order_crossover partially_matched_crossover edge_recombination_crossover
    mutation = inverse_mutation, #gaperm_swMutation  # Mutation function swap_mutation inverse_mutation scramble_mutation insert_mutation
    elitism = elitism
  )

summary(ga_result)
```

